%!TEX root = ../../report.tex

%Rope
\subsubsection{Rope}

 is a Python refactoring tool written in Python, which works like a Python library.
In order to make it easier to create refactoring operations Rope assumes that a Python program only has assignments and functions calls. %(can this be a bad thing?)
The tool can easily get information about the assignments. 
However for functions calls it is necessary to have other approaches in order to obtain the necessary information. 

Rope uses a Static Object Analysis which analyses the modules or scope to get information about functions. 
Rope only analyses the scopes when they change and it only analyses the modules when asked by the user, because this approach is time consuming. 

The other approach is the Dynamic Object Analysis that starts working when a module is running. 
Then, Rope collects information about parameters passed to and returned from functions in order to get all the information needed. 
The main problem is that this approach is slow while collecting information, but not while accessing the information.

Rope stores the information collected by the analysis in a database. 
If Rope needs the information and there is nothing on the database the Static object inference starts trying to infer the object information.

Rope uses an AST in order to store the syntax information about the programs.

%It has simple refactoring operations such as, Rename, Extract method/local variable, Move, inline, Introduce factory, Change method signature, Transform module to package, Encapsulate field and Replace method with method object.

%And it also can: Extract similar statements in extract refactorings, fix imports when needed, preview refactorings, undo/redo refactorings, interrupt refactorings, perform cross-project refactorings, handle basic implicit interfaces in rename and change signature.

%And helps IDE's with:

 %   Auto-completion
  %  Finding definition location
   % Getting pydoc
    %Finding occurrences
    %Organizing imports (removing unused and duplicate imports and sorting them)
    %Generating python elements

 %(the PyRefactor a plugin for sublime text 3 that uses Rope https://packagecontrol.io/packages/PyRefactor)

 %It has a module that tries to support build-in types and functions.
