\section{Architecture}
%Refactoring Tool work flow: (add a brief explanation of what does)

%FIXME The Chosen IDE here??

In this section we present the architecture of our refactoring tool.
Although the tool can work with different programming languages, its
main focus is, at this moment, the Racket programming language and,
more specifically, the DrRacket IDE.

Racket is a language designed to support meta-programming and, in
fact, most of the syntax forms of the language are macro-expanded into
combinations of simpler forms.  This has the important consequence
that programs can be analyzed either in their original form or in
their expanded form.

In order to create correct refactoring operations, the refactoring
tool uses two sources of information, the def-use relations and the
AST of the program.  The def-use relations represent the links between
definition of an identifier and its usage.  In the DrRacket IDE these
relations are visually represented as arrows that point from a
definition to its use.  The opposite relation, the use-def relation is
also visually represented as an arrow from the use of an identifier to
its definition.  The AST is the abstract syntax tree of the program
which, in the case of the Racket language, is represented by a list of
syntax-objecs.  Given that Racket
%In Racket everything is a syntax-object and therefore accessing the list
%of syntax-objects has all the information that a normal AST provides. %FIXME incorrect


\begin{figure}
\centering
\tikzstyle{startstop} = [rectangle,  minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=red!30] %start
\tikzstyle{process} = [rectangle,  minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=orange!30] %process
\tikzstyle{process2} = [rectangle,  minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
\tikzstyle{process3} = [rectangle,  minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
\tikzstyle{process4} = [rectangle,  minimum width=3cm, minimum height=2.5cm, draw=black, fill=blue!30]
\tikzstyle{arrow} = [thick,->,>=stealth]
\begin{tikzpicture}[node distance=2cm]
\node (pro1) [process] {Reader};
\node (pro2) [process, below of=pro1] {Expander};
\node (pro3) [process, below of=pro2] {def-use relations};
\node (pro5) [process4, right of=pro2, xshift=1cm, align=left] { \\[2em] Transformer};
\node (pro4) [process3, xshift=-1.03cm, yshift=0.75cm] at (pro5.east) {Code
 Walker};
\node (pro6) [process2, right of=pro5, xshift=1cm] {Writer};
\draw [arrow] (pro1) -- (pro2);
\draw [arrow] (pro2) -- (pro3);
\draw [arrow] (pro1) -- (pro5);
\draw [arrow] (pro2) -- (pro5);
\draw [arrow] (pro3) -| (pro5);
%\draw [arrow] (pro4) -- (pro5);
%\draw [arrow] (pro5) -- node[anchor=east] {use} (pro4);
\draw [arrow] (pro5) -- (pro6);
\end{tikzpicture}
\caption{Main modules and information flow between modules. Unlabeled arrows represent information
flow between modules.}
\label{flow-chart}
\end{figure}

Figure~\ref{flow-chart} summarizes the workflow of the refactoring
tool, where the Reader produces the non expanded AST of the program
while the Expander expands the AST produced by the Reader.  In order
to produce the def-use relations it is necessary to use the expanded
AST produced by the Expander because it has the correct dependency
information.  The Transformer uses the Code Walker to parse the ASTs
and the information of the def-use relations to correctly perform the
refactoring operations.  Then it goes to the Writing module to produce
the output in DrRacket's definitions pane.


\subsection{Syntax Expressions}
%intro and explain the importance of the S-Exp
The syntax-object list represents the AST, which provides information
about the structure of the program.  The syntax-object list is already
being produced and used by the Racket language and, in DrRacket, in
order to provide error information to the user.  DrRacket already
provides functions which computes the program's syntax-object list and
uses some of those functions in the Background Check Syntax and in the
Check Syntax button callback.


\subsubsection{Syntax Expression tree forms}
DrRacket provides functions to compute the syntax-object list in two
different formats.  One format is the expanded program, which computes
the program with all the macros expanded.  The other format is the
non-expanded program and computes the program with the macros
unexpanded.

The expanded program has the macros expanded and the identifier
information correctly computed.  However, it is harder to extract the
relevant information when compared with the non expanded program.

For example, the following program is represented in the expanded
form, and in the non expanded form.

\begin{lstlisting}[basicstyle=\ttfamily, caption=Original Code]
(and alpha beta)
\end{lstlisting}

\begin{lstlisting}[basicstyle=\ttfamily, caption=Expanded program]
#<syntax:2:0
 (#%app call-with-values
 (lambda ()
    (if alpha beta (quote #f)))
  print-values)>
\end{lstlisting}

\begin{lstlisting}[basicstyle=\ttfamily, caption=Non-expanded program]
#<syntax:2:0 (and alpha beta)>
\end{lstlisting}

Note that the expanded program transforms the {\tt and}, {\tt or},
{\tt when}, and {\tt unless} forms into {\tt if}s which makes
refactoring operations harder to implement.

Racket adds internal representation information to the expanded
program which for most refactoring operations is not necessary.  In
addition, the expanded program has a format that is likely to change
in the future.  Racket is an evolving language and the expanded form
is a low level and internal form of representation of the program.
However, the expanded program has important information regarding the
binding information that is not available in the non-expanded form,
and this information might be useful, e.g., to detect if two
identifiers refer to the same binding.  Additionally, we do not
consider macro definitions as part of the code that needs to be
refactored, since the refactoring tool is targeted at unexperienced
programmers and these programmers typically do not define macros.

Taking the previous discussion into consideration, it becomes clear
that it is desirable to use the non expanded form for the
refactoring %all things considered
operations whenever possible and use the expanded form only when
needed.%   Nevertheless, if we intended to create a tool that gives
% support to refactoring macros we would need to use the expanded
% program.  However, there are no guarantees that would be enough to
% ensure the correctness of such refactoring operations due to the
% reflection capabilities of Racket.


\subsection{Def-use relations}
Def-use relations hold important information needed in order to
produce correct refactoring operations.  They can be used to check
whether there will be a duplicated name or to compute the arguments of
a function that is going to be extracted.

% DrRacket already uses the def-use-relations in the system and they are
% visually %TODO they are or it is?
% represented by arrows in the GUI.
The def-use-relations is computed by the % online-
compiler that runs in the background.  However, it is only computed
when a program is syntactically correct.

\subsection{Code-walker}
The code-walker is used to parse the syntax tree represented by a
syntax element that is a list of syntax-object in Racket.  A
syntax-object can contain either a symbol, a syntax-pair, a datum
(number, boolean or string), or an empty list.  While a syntax-pair is
a pair containing a syntax object as it first element and either a
syntax pair, a syntax element or an empty list as the second argument.
Each syntax-object has information about the line where they are
defined and this information is used to search for the correct
elements.


Most of the time, the code-walker is used to search for a specific
syntax element and the location information contained in the
syntax-object is used to skip the syntax blocks that are before the
syntax element wanted in the first place.

The Code-walker is a core part of the refactoring tool, ensuring that
the selected syntax is correctly fed to the refactoring
operations. %actually?


\subsection{Pretty-printer}
Producing correct output is an important part of the refactoring tool.
It is necessary to be careful to produce indented code and we decided
to use a pretty-printer that is already available in the Racket
language.  However, it should be noted that this pretty-printer does
not follow some of the Racket style convention, such as {\tt cond}
clauses surrounded by square brackets. This is not considered a
problem because Racket supports both representations.  One possible
solution is to use a different pretty-printer in order to keep the
language conventions.


\subsection{Comments preservation}
Preserving the comment information after a refactoring transformation
is an important task of the refactoring tool. If the comment in
determined place of the program changes its location, affecting a
different part of the program, it could confuse the programmer.
However, comment preservation is not implemented yet, making it a
limitation of this prototype.

One possible solution is to modify the syntax reader and add a comment
node to the AST.  While the new node will not be used during
refactoring transformations it is used during the output part of the
refactoring operation, preserving the comment with the correct syntax
expression.


\subsection{Syntax-Parse}

The Syntax-Parse\cite{tobin2011languages} function provided by Racket
is very useful for the refactoring operations. It provides a wide
range of options to help matching the correct syntax,
using %fine tunning
backtracking to allow have several rules to be matched in the same
syntax parser, which helps to create more sophisticated rules.

\section{Refactoring operations}
In this section we explain some of the  more relevant refactoring operations and
some limitations of the refactoring tool.

\subsection{Semantic problems}
There are some well-known semantic problems that might occur after
doing a refactoring operation.  One of them occurs in the refactoring
operation that removes redundant {\tt and}s in numeric comparisons.
Although rarely known by beginner programmers, in Racket, numeric
comparisons support more than two arguments, as in \verb|(< 0 ?x 9)|,
meaning the same as %
\verb|(and (< 0 ?x) (< ?x 9)|, where, we use the notation \verb|?x| to
represent an expression. Thus, it is natural to think about a
refactoring operation that eliminates the \verb|and|.  However, when
the \verb|?x| expression somehow produces side-effects, the
refactoring operation will change the meaning of the program.
% \begin{lstlisting}[basicstyle=\ttfamily, caption=And example]
%   (define (foo arg)
%     (displayln "foo")
%     arg)
%   (and (< 1 (foo 2)) (< (foo 2) 3))
% \end{lstlisting}
% The refactoring transforms the code into this:
% \begin{lstlisting}[basicstyle=\ttfamily, caption=Refactored code]
%   (< 1 (foo 2) 3)
% \end{lstlisting}


% Instead of applying the side effect that is displaying the the string "foo"
%  twice it will only display it once. Therefore changing the meaning of the program.

Despite this problem, we support this refactoring operation because,
in the vast majority of the cases, there are no side-effects being
done in the middle of numerical comparisons and, therefore, the
refactoring operation does not change the meaning of the program.
% Furthermore, the possible solution would limit excessively this
% refactoring operation.  %enormously
% Considering Racket's reflection capabilities we would only apply this
% refactoring operation safely when the arguments of the {\tt <}
% expressions, in this case the numbers 1, 2, 3, and the function {\to
%   the foo} were datums (number, boolean or string).


Another example of a semantic problem occurs when refactoring the following {\tt if}
expression.
\begin{lstlisting}[basicstyle=\ttfamily, caption=Code sample]
(if ?x
    (begin ?y ...)
    #f)
\end{lstlisting}
There are two different refactoring transformations possible:
\begin{lstlisting}[basicstyle=\ttfamily, caption=Refactoring option 1]
(when ?x
  ?y ...)
\end{lstlisting}

\begin{lstlisting}[basicstyle=\ttfamily, caption=Refactoring option 2]
(and ?x (begin ?y ...))
\end{lstlisting}

Note that the first refactoring option changes the meaning of the
program, because if the test expression, in this case {\tt ?x}, is
false, the result of the {\tt when} expression is {\tt \#<void>}.
However, the programmer may still want to choose the first refactoring
option if the return value when the {\tt ?x} is false is not
important.% ,
% there are two possible refactoring operations applicable.
% For example, if a programmer is creating a predicate may
% choose the {\tt and} version, whereas if the programmer is using another control structure
% and do not care for the result of the expression may prefer the {\tt when} version.

\subsection{Extract Function}
Extract function is an important refactoring operation that every
refactoring tool should have.  In order to extract a function it is
necessary to compute the arguments needed to the correct use of the
function.  While giving the name to a function seams quite
straightforward, it is necessary to check for name duplication in
order to produce a correct refactoring as having two identifiers with
the same name in the same scope produces an incorrect program.  After
the previous checks, it is straightforward to compute the function
body and replacing the original expression by the function call.

However, the refactoring raises the problem where should the function
be extracted to. A function can not be defined inside an expression,
but it can be defined at the top-level or at any other level that is
accessible from the current level.

As an example, consider the following program:
\begin{lstlisting}[basicstyle=\ttfamily, caption=Extract function levels]
;;top-level
(define (level-0)
  (define (level-1)
    (define (level-2)
      (+ 1 2))
    (level-2))
  (level-1))
\end{lstlisting}

When extracting the {\tt (+ 1 2)} to a function where should it be
defined?  Top-level, Level-0, level-1, or in the current level, the
level-2?  The fact is that is extremely difficult to know the answer
to this question because it depends on what the user is doing and the
user intent.  Accordingly, we decided that the best solution is to let
the user decide where the user wants the function defined.


\subsubsection{Computing the arguments}

In order to compute the function call arguments we have to know in
which scope the variables are being defined, in other words, if the
variables are defined inside or outside the extracted
function. %FIXME it is weird
The variables defined outside the function to be extracted are
candidates to be the arguments %FIXME IMPROVE
of that function.  However, imported variables, whether from the
language base or from other libraries do not have to be passed as
arguments. To solve this problem, we considered two possible
solutions:

\begin{itemize}
\item Def-use relations + Text information
\item Def-use relations + AST
\end{itemize}

The first approach is simpler to implement and more direct than the
second one.  However, it is less tolerant to future changes and to
errors.  The second one combines the def-use relations information
with the syntax information to check whether it is imported from the
language or from other library.

We choose the second approach in order to provide a more stable
solution to correctly compute the arguments of the new function.

\subsection{Let to Define Function}
%there are several let forms, we want to focus on let* let and named let.
A {\tt let} expression is very similar to a function, which may led the user
to mistakenly use one instead of the other.
Therefore we decided to provide a refactoring operation that would make such transition simpler.


There are several let forms, but since we want to explore the similarity between the {\tt let} and the function
we are going to focus in the ones that are more similar to a function, namely the {\tt let} and the {\tt named let}.

There are some differences between them, the {\tt named let} can be directly mapped to a named function, using the {\tt define} keyword,
whereas the {\tt let} can only be directly mapped to an anonymous function, {\tt lambda}.
We decided to focus first in the transformation of a {\tt named let} to a function.

However, this refactoring operation which transforms a {\tt named let} into a {\tt define} function,
 could have syntax problems since a {\tt let} form can be used in expressions, but the {\tt define} can not.
In the vast majority of cases this refactoring is correct, however, when a {\tt named let} is used in an expression
it transforms the program in an incorrect one.
e.g.
\begin{lstlisting}[basicstyle=\ttfamily, caption=Let in an expression]
(and (let xpto ((a 1)) (< a 2)) (< b c))
\end{lstlisting}
Modifying this {\tt named let} into a {\tt define} would raise a syntax error since a
{\tt define} could not be used in an expression context.
Encapsulate the {\tt define} with the {\tt local} keyword, which is an expression like
the {\tt named let}, can solve the problem.
However, the {\tt local} keyword is not used very often and might confuse the users.
Therefore we decided keep the refactoring operation without the local keyword that works for
most of the cases.


\subsection{Wide-Scope Replacement}
The Wide-Scope replacement extends the extract function functionality by replacing
the duplicated of the extracted code with the function call of the extracted function.


The Wide-Scope replacement refactoring operation searches for the code that is duplicated of the extracted
function and then replaces it for the call of the
extracted function and it is divided in two steps:
\begin{itemize}
\item Detect duplicated code
\item Replace the duplicated code
\end{itemize}

Replacing the duplicated code is the easy part, however the tool might has to compute
the arguments for the duplicated code itself.

Correctly detecting duplicated code is a key part for the correctness of this refactoring.
Even the simplest form of duplicated code detection, where it only detects duplicated code
when the code is exactly equal, may have some problems regarding the binding information.
For example, if the duplicated code is inside a {\tt let} that changes some bindings that must
be taken into consideration.
In order to solve the binding problem we can use functions already provided in Racket.
However, that does not work if we use the program in the not expanded
form to do the binding comparisons because there is not enough information for those bindings to work.
Therefore, in order to compute the correct bindings, it is necessary to use the expanded form
of the program.

The naive solution is to use the expanded program to detect the duplicated
 code and then use this information to do the replacing of the duplicated code.
However, when expanding the program Racket adds necessary internal information to
run the program itself that are not visible for the user.
While this does not change the detecting of the duplicated code, it adds unnecessary information
that would have to be removed.
In order to solve this in a simple way we can use the expanded code to detect
the correctly duplicated code and use the non expand program
to compute which code will be replaced.

However, this duplicated code detection is a quadratic algorithm which might
have some performance problems for bigger programs.

%Detecting duplicated code can be added to the automatic detection of possible refactoring operations to be applied. %FIXME is this already done?
%Notifying the users of a possible extract function operation if there is duplicated code is a rather
%useful because sometimes it might be difficulty for the user to remember if a
%piece of code is duplicated or not.
