\lstdefinestyle{racket}{
  language=Lisp,
  basicstyle=\tt\scriptsize,
  columns=fullflexible,
  commentstyle=\color{Gray},
  keywordstyle=\color{Blue},
  identifierstyle=\color{Blue},
  stringstyle=\color{Green},
  numbers=left,
  numberstyle=\small\tt,
  frame=lines,
}

\lstdefinestyle{racket}{
  language=Lisp,
  basicstyle=\tt\scriptsize,
  columns=fullflexible,
  commentstyle=\color{gray},
  keywordstyle=\color{blue},
  identifierstyle=\color{black},
  stringstyle=\color{green},
  numbers=left,
  numberstyle=\small\tt,
  frame=none,
  morekeywords={define,if,else,cond},
  literate={% replace strings with symbols
    {lambda}{{$\lambda$}}{1}
  },
}
\lstdefinestyle{python}{
  language=Python,
  basicstyle=\tt\scriptsize,
  columns=fullflexible,
  commentstyle=\color{Gray},
  keywordstyle=\color{Orange},
  identifierstyle=\color{Blue},
  stringstyle=\color{Green},
  numbers=left,
  numberstyle=\small\tt,
  frame=lines,
}

\section{Evaluation}
In this section we present some code examples written by beginner programmers
in their final project of an introductory course and their possible improvements
using the refactoring operations available in our refactoring tool.
The examples show the usage of some of the refactoring operations previous presented
and here is explained the motivation for their existence.

\lstset{style=racket}


The first example is a very typical error made beginner programmers.
\begin{lstlisting}
(if (>= n_plays 35)
  #t
  #f)
\end{lstlisting}
It is rather a simple refactoring operation, but nevertheless it improves the code.
%if after , caption=Refactored If expression
\begin{lstlisting}
(>= n_plays 35)
\end{lstlisting}

%and before , caption=And expression
The next example is related with the conditional expressions, namely the {\tt and}
or {\tt or} expressions. We decided to choose the {\tt and} expression to exemplify a
rather typical usage of this expression.
\begin{lstlisting}
(and
  (and
   (eq? #t (correct-moviment? player play))
   (eq? #t (player-piece? player play)))
  (and
   (eq? #t (empty-destination? play))
   (eq? #t (empty-start? play))))
\end{lstlisting}


%and after
Transforming the code by removing the redundant {\tt and} expression makes the code
cleaner and simpler to understand.
\begin{lstlisting}
(and (eq? #t (correct-moviment? player play))
     (eq? #t (player-piece? player play))
     (eq? #t (empty-destination? play))
     (eq? #t (empty-start? play)))
\end{lstlisting}

However, this code can still be improved, the {\tt (eq? \#t ?x)} is a redundant way
of simple writing {\tt ?x}.
\begin{lstlisting}
(and (correct-moviment? player play)
     (player-piece? player play)
     (empty-destination? play)
     (empty-start? play))
\end{lstlisting}

While a student is still learning is common to forget whether or not a
sequence of expressions need to be wrapped in a {\tt begin} form.  The
{\tt when}, {\tt cond} and {\tt let} expressions have a implicit {\tt
  begin} and as a result it is not necessary to add the {\tt begin}
expression.  Moreover, sometimes they still kept the {\tt begin}
keyword because they often use a trial and error approach in writing
code.  Our refactoring tool checks for those mistakes and corrects
them.
%let with begin before, Chose a better example
\begin{lstlisting}
(if (odd? line-value)
    (let ((internal-column (sub1 (/ column 2))))
      (begin
        (if (integer? internal-column)
            internal-column
            #f)))
    (let ((internal-column (/ (sub1 column) 2)))
      (begin
        (if (integer? internal-column)
            internal-column
            #f))))
\end{lstlisting}
%let with begin after
This is a simple refactoring operation and it does not have a big
impact, however it makes the code clear and helps the beginner
programmer to learn that a {\tt let} does not need a {\tt begin}.
\begin{lstlisting}
(if (odd? line-value)
    (let ((internal-column (sub1 (/ column 2))))
      (and (integer? internal-column)
           internal-column))
    (let ((internal-column (/ (sub1 column) 2)))
      (and (integer? internal-column)
           internal-column)))
\end{lstlisting}


The next example shows a nested {\tt if}. Nested {\tt if}s are
 difficult to understand, error prone, and a debugging nightmare.
\begin{lstlisting}
(define (search-aux? board line column piece)
  (if (> column 8)
    #f
    (if (= line 1)
      (if (eq? (house-board board 1 column) piece)
        #t
        (search-aux? board line (+ 2 column) piece))
      (if (= line 2)
        (if (eq? (house-board board 2 column) piece)
          #t
          (search-aux? board line (+ 2 column) piece))
        (if (= line 3)
          (if (eq? (house-board board 3 column) piece)
            #t
            (search-aux? board line (+ 2 column) piece))
          (if (= line 4)
            (if (eq? (house-board board 4 column) piece)
              #t
              (search-aux? board line (+ 2 column) piece))
            (if (= line 5)
              (if (eq? (house-board board 5 column) piece)
                #t
                (search-aux? board line (+ 2 column) piece))
              (if (= line 6)
                (if (eq? (house-board board 6 column) piece)
                  #t
                  (search-aux? board line (+ 2 column) piece))
                (if (= line 7)
                  (if (eq? (house-board board 7 column) piece)
                    #t
                    (search-aux? board line (+ 2 column) piece))
                  (if (= line 8)
                    (if (eq? (house-board board 8 column) piece)
                      #t
                     (search-aux? board line (+ 2 column) piece))
                    null))))))))))
\end{lstlisting}

It is much simpler to have a {\tt cond} expression instead of the nested {\tt if}.
In addition, every true branch of this nested if contains {\tt if} expressions that are
or expressions and by refactoring those {\tt if} expressions to {\tt or}s it makes the code
simpler to understand.


\begin{lstlisting}
(define (search-aux? board line column piece)
  (cond [(> column 8) #f]
        [(= line 1)
         (or (eq? (house-board board 1 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [(= line 2)
         (or (eq? (house-board board 2 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [(= line 3)
         (or (eq? (house-board board 3 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [(= line 4)
         (or (eq? (house-board board 4 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [(= line 5)
         (or (eq? (house-board board 5 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [(= line 6)
         (or (eq? (house-board board 6 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [(= line 7)
         (or (eq? (house-board board 7 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [(= line 8)
         (or (eq? (house-board board 8 column) piece)
             (search-aux? board line (+ 2 column) piece))]
        [else null]))
\end{lstlisting}

However, this code could still be further improved by refactoring it into a {\tt case}. \\

The examples presented above appear repeatedly in almost every code submission of this final project supporting the need to
proper support to beginner programmers.
\begin{table}[]
\centering
\caption{Refactoring Operations}
\label{tab:ref-ops}
\begin{tabular}{c|c|c|c|c|c|c}
Code \#                                                     & 1     & 2      & 3     & 4     & 5      & Total \\ \hline
Initial LOC                                                 & 582   & 424    & 332   & 1328  & 810    & 3476  \\ \hline
Final LOC                                                   & 545   & 373    & 300   & 1259  & 701    & 3178  \\ \hline
Difference                                                  & 37    & 51     & 32    & 69    & 109    & 298   \\ \hline
Percentage                                                  & -6.36 & -12.03 & -9.64 & -5.19 & -13.46 & -9.34 \\ \hline
Remove Begin                                                & 11    & 4      & 6     & 9     & 5      & 35    \\ \hline
If to When                                                  & 4     & 0      & 0     & 0     & 0      & 4     \\ \hline
If to And                                                   & 3     & 1      & 0     & 0     & 0      & 4     \\ \hline
If to Or                                                    & 6     & 6      & 1     & 13    & 20     & 46    \\ \hline
Remove If                                                   & 0     & 3      & 7     & 6     & 3      & 19    \\ \hline
Remove And                                                  & 0     & 2      & 0     & 4     & 0      & 6     \\ \hline
Remove Eq                                                   & 0     & 4      & 0     & 0     & 0      & 4     \\ \hline
\begin{tabular}[c]{@{}c@{}}Extract \\ Function\end{tabular} & 0     & 3      & 0     & 0     & 4      & 7     \\ \hline
If to Cond                                                  & 0     & 0      & 0     & 2     & 3      & 5
\end{tabular}
\end{table}


As seen in Table~\ref{tab:ref-ops} the average reduction in LOC is 9.34\% which
shows how useful are these refactoring operations. It also shows how many
refactoring operations were applied.
This tool is not only for beginners, during the development of the tool we already
used some of the refactoring operations, namely the extract function, in order to
improve the structure of the code.

%if the users had a refactoring tool that highlighted their code in areas that
%could be improved, they probably would not have this kind of code.

\subsection{Refactoring Python}
Python is being promoted as a good replacement for Scheme and Racket in science introductory courses.
It is an high-level, dynamically typed programming language and it supports the functional, imperative and object
oriented paradigms.
Using this architecture and the capabilities offered by Racket and DrRacket with an implementation of Python for Racket\cite{ramos2014implementation} \cite{ramos2014reaching}
we can also have refactoring operations in Python. \\
%focus in the functional paradigm of Python
Using Racket's syntax-objects to represent Python like a meta-language like it was used in FAMIX\cite{tichelaar2000meta}.
It is possible use the same structure to parse and analyze the code used for the refactoring operations in Racket.

However, there are some limitations regarding the refactoring operations in Python.
Since Python is a statement base language instead of expression base, it raises
some problems regarding the possibility of some refactoring operations.

Example of removing an If expression:
\lstset{style=python}
\begin{lstlisting}
True if (alpha < beta) else False
\end{lstlisting}

\begin{lstlisting}
(alpha < beta)
\end{lstlisting}

The next one shows the extract function:
%\begin{lstlisting}
%def complexFunction(a, b, c):
%    a + b + c
%    c + b - a
%\end{lstlisting}

%\begin{lstlisting}
%def simpleAddition( a,b,c):
%    return a + b + c
%def complexFunction(a, b, c):
%    simpleAddition( a,b,c)
%    c + b - a
%\end{lstlisting}
\begin{lstlisting}
def mandelbrot(iterations, c):
    z = 0+0j
    for i in range(iterations+1):
        if abs(z) > 2:
            return i
        z = z*z + c
    return i+1
\end{lstlisting}

\begin{lstlisting}
def computeZ(z, c):
    return z*z + c

def mandelbrot(iterations, c):
    z = 0+0j
    for i in range(iterations+1):
        if abs(z) > 2:
            return i
        z = computeZ( z,c)
    return i+1
\end{lstlisting}

This refactoring operations for Python are just only a prototype and it is a work
in progress.
\section{Conclusion}
A refactoring tool designed for beginner programmers would benefit them
by providing a tool to restructure the programs and improve what more knowledgeable programmers call ``poor style,''
``bad smells,'' etc.
In order to help those users it is necessary to be usable from a pedagogical IDE,
to inform the programmer of the presence of the typical mistakes made by beginners, and
to correctly apply refactoring operations preserving semantics.

Our solution tries to help those users to improve their programs by using the AST of
the program and the def-use-relations to create refactoring operations that do not
change the program's semantics. This structure is then used to analyze the code
to detect typical mistakes using automatic suggestions and correct them using the
refactoring operations provided.


There are still some improvements that we consider important for the user.
Firstly, the detection of duplicated code is still very naive and improving to understand if
two variables represent the same even if the names are different or even if the
 order of some commutative expressions is not the same would make a huge improvement
 on the automatic suggestion.
Then it is possible to improve the automatic suggestion of refactoring operations by
having different colors for different types of refactoring operations.
With a lower intensity for low "priority" refactoring operations and a high intensity
for higher "priority". Thus giving the user a better knowledge of what is a better
way to solve a problem or what is a strongly recommendation to change the code.
Lastly it would be interesting to detect when a developer is refactoring in order to help the developer finish the
refactoring by doing it automatically \cite{ge2012reconciling}.
